---
title: "Homework 1"
author: "Put Your Name Here"
date: 'Put Date Here'
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use the code chunk below to load any R libraries that you use in your script. I've added `dplyr` as a start.

```{r, message = FALSE}
library(dplyr)
```

<br>

# Part 1 - Mario Kart Example

1. Install the `moderndive` package on your computer if you haven't done so yet. Then add code above to load the package in this document. Explain in your own words below why we do not want to include the code to install the package in this document.

**Answer:** REPLACE THIS WITH YOUR ANSWER

<br>



2. Write code below to bring up the help menu for the `mario_kart_auction` data set from the `moderndive` package. Then, tell me how many observations are in the data and give an example of at least two quantitative variables and at least two categorical variables.

```{r}
# Put your code below to bring up the help menu

```

**Number of observations:**

**Quantitative variables:**

**Categorical variables:**

<br>



3. Use a function to display the number of games that are new vs. used. Then add another line of code that displays the proportion of games that are new vs. used. After the code, tell me the percentage of games that were used.

```{r}
# Put your code below to explore the data

```

**Percentage of Used Games:**

<br>



4. Write code that creates an object called `mario_sample` that includes just the first 15 observations in the data. Add another line of code that prints only the overall auction price (`total_pr`) for those 15 observations. In your answer below, use in-line R code (Hint: we did this in Activity 02) to tell someone the highest overall price among your sample of 15 auctions.

```{r}
# Put your code below

```

**Sentence with in-line R code:** 

<br>



5. In the previous question, what type of object is `mario_sample`? What type of object is the 15 overall auction prices?

**Answer:**

<br>



6. Create a data frame called `mario_summary` that summarizes the data by calculating the following:

  - number of auctions (called `auctions`)
  - mean shipping price (called `avg_ship`)
  - longest duration (called `max_length`)
  - standard deviation of number of bids (called `sd_bids`)

After the data frame is created, print the data frame so the values are visible.

```{r}
# Put your code below

```

<br>



7. Take your code from the previous question and make it so that a separate summary is calculated for new games and for used games. In other words, your `mario_summary` data frame will now have two rows (new and used). After the data frame is created, print the data frame so the values are visible. Is there more variation for number of bids among new games or used games? Briefly explain how you know.

```{r}
# Put your code below

```

**Answer:**

<br>



8. Take your `mario_summary` object and display it below using the `kableExtra` package to format it nicely for this document. Add an option to this code chunk so that your code is **not** displayed in the knitted document. (Note: You should also add code at the top of the document to load `kableExtra` where other packages are loaded.)

```{r}
# Put your code below

```

<br>



# Part 2 - Miscellaneous

The questions in this section do not directly apply to the previous questions.

9. Is each of the following an argument, function, object, or operator? If the item is an object, tell me whether it is a scalar, vector, matrix, or data frame. Put your answer after the colon for each piece of code.

  - `c()`: 
  - `$`: 
  - the `n` in `head(__, n = __)`: 
  - `mario_kart_auction$n_bids`:

<br>



10. Write code to create a vector called `spring_25` that includes the courses you're taking this semester (e.g. STS2300).

```{r}
# Put your code below

```

<br>



11. Answer one (or both) of the following questions. Your answer should be at least 3 - 4 sentences long and demonstrate some level of reflection (i.e., not just a few words). Be sure to answer the full question as well.

  - What's something from Notes 01 or 02 that isn't entirely clear to you *and* what do you think would help to solidify this concept for you? Be specific.
  - What's something that was surprising to you as you first learned about R in this class (or built on your previous R knowledge)? What made it surprising?
  
**Answer:** 
